---
title: "Lab 9"
author: "Desiree Smith"
date: "11/9/2021"
output: 
 html_document:
    highlight: zenburn
    toc: TRUE
    toc_float: 
      collapsed: FALSE
---

```{r Walkthrough , include=FALSE}
require(here)
catrate = read.csv(here("data", "catrate.csv"))
head(catrate)


#binomial test
n_success = sum(catrate$success)
n_years = sum(catrate$years)
binom.test(n_success, n_years)

binom.test(n_success, n_years, p = 5/7) 
binom.test(
  n_success,
  n_years,
  p = 5/7,
  alternative='less')

veg = read.csv(here("data", "vegdata.csv"))
head(veg)

#boxplot
boxplot(pine ~ treatment, data = veg)

#variance test
var.test(
  pine ~ treatment,
  data = veg,
  subset = treatment %in% c('control','clipped'))

#shapiro test 
shapiro.test(veg$pine[veg$treatment=="control"])

shapiro.test(veg$pine[veg$treatment=="clipped"])

fligner.test(
  pine ~ treatment,
  data = veg,
  subset = treatment %in% c('control','clipped'))

bartlett.test(pine ~ treatment, data=veg)

fligner.test(pine ~ treatment, data = veg)

#t test
t.test(pine~treatment,data=veg,subset=treatment %in% c('control','clipped'), conf.int=TRUE)


```

```{r walkthrough, include=FALSE }

#Wilcox test
wilcox.test(pine~treatment,data=veg,subset=treatment %in% c('control','clipped'), conf.int=TRUE)

#create separate vectors for the “control” observations and “clipped” observations
control = veg$pine[veg$treatment=='control']
clipped = veg$pine[veg$treatment=='clipped']

#paired t-test
t.test(control, clipped, paired=TRUE)

wilcox.test(control, clipped, paired=TRUE)

disp = read.csv(here("data", "dispersal.csv"))

disp

plot(disp$disp.rate.ftb, disp$disp.rate.eb)

#testing the significance of the correlation
cor.test(
  disp$disp.rate.ftb,
  disp$disp.rate.eb,
  use='complete.obs')

cor.test(
  disp$disp.rate.ftb,
  disp$disp.rate.eb,
  use='complete.obs',
  method='spearman')

#sample of dispersal rates
plot(
  ecdf(disp$disp.rate.ftb),
  verticals=TRUE)

#adult dispersal rate 
plot(
  ecdf(disp$disp.rate.ftb),
  verticals=TRUE)
plot(
  ecdf(disp$disp.rate.eb),
  verticals=TRUE,
  lty=3,
  add=TRUE)
```

```{r walkthrough , include=FALSE}
ks.test(disp$disp.rate.ftb,disp$disp.rate.eb)

#This is a simple binomial proportions test, which we can easily do in R by specifying two vectors:the number of mortalities for females and males c(4,16)the total number of female and male candidates: c(40,250)

prop.test(c(4,16),c(40,250))

#Fisher's exact test
fisher.test

#bird habitat data
birds   = read.csv(here("data", "bird.sta.csv"))
hab     = read.csv(here("data", "hab.sta.csv"))
birdhab = merge(birds, hab, by=c("basin", "sub", "sta"))

# Create a contingency table for edge/interior and brown creeper presence/absence
table(birdhab$s.edge, birdhab$BRCR > 0)

# set the presence to be in the first column
br_creeper_table = table(birdhab$s.edge, birdhab$BRCR > 0)[, 2:1]

owls = matrix(c(16, 9, 4, 11), nrow=2)
rownames(owls) = c("present", "absent")
colnames(owls) = c("old", "young")
chisq.test(owls)


```


### Question 1
State the null hypothesis of the Chi-square test.

The null hypothesis of the chi- square test is that there is no significant relationship for the data. 


### Question 2
Consider the results of your test and explain whether you think that Brown Creepers show a significant habitat preference.

Based on the results of test you can see that the brown creepers do show significant a significant relationship. The p-value is very small so we will reject the null.

```{r}
owls = matrix(c(16, 9, 4, 11), nrow=2)
rownames(owls) = c("present", "absent")
colnames(owls) = c("old", "young")
chisq.test(br_creeper_table)
```

### Question 3 -5

```{r}
require(palmerpenguins)
penguins

fit_species =
  lm( formula = body_mass_g
    ~ species,
    data = penguins)
fit_species

fit_sex = 
  lm(
    formula = body_mass_g ~ sex,
    data = penguins)
fit_sex

fit_both = 
  lm(
    formula = body_mass_g ~ sex * species,
    data = penguins)
fit_both

```

### Question 6
 Include a conditional boxplot corresponding to your fit_species model.

```{r}
boxplot(body_mass_g ~ species,
    data = penguins, ylab = "body mass (g)",col= "lightskyblue1"
             
             )


```

### Question 7
Include a conditional boxplot corresponding to your fit_sex model.

```{r}
boxplot(body_mass_g ~ sex,
    data = penguins, ylab = "body mass (g)",col= "peachpuff1")
             
```




### Question 8
 Include a conditional boxplot corresponding to your fit_species model.
 
```{r}
#Example
boxplot(body_mass_g ~ sex * species,
    data = penguins, ylab = "body mass (g)",col= "lavender",
   names = c("Adelie\n female",
             "Adelie\n male",
             "Chinstrap\n female",
             "Chinstrap\n male",
             "Gentoo\n female",
             "Gentoo\n male"
             
             ))


```

### Question 9
Based on the shapes of the boxes, which of the models (if any) do you think may have problems fulfilling the heterogeneity assumption?

The model that includes penguin body mass of species does not fulfill the heterogeneity assumption. The shapes of the boxes vary the most out of all of the models. 

### Question 10
State the null hypothesis of the Bartlett test.

The Null hypothesis of the Bartlett test is that the all of the variances are equal to each other.

### Question 11
 What was the p-value from the Bartlett test of homogeneity for observations grouped by species?
 
The p-value for the test is 0.050.
```{r}
bartlett.test(body_mass_g ~ species, data= penguins)
```

### Question 12
What was the p-value from the Bartlett test of homogeneity for observations grouped by sex?

The p-value is 0.032.
```{r}
bartlett.test(body_mass_g ~ sex, data= penguins)
```

### Question 13
What was the p-value from the Bartlett test of homogeneity for observations grouped by both factors?

The p-value of Bartlett test 0.1741. It is not very significant.

```{r}
dat_groups = aggregate(
  body_mass_g ~ species * sex,
  data = penguins,
  FUN = c)
str(dat_groups)

bartlett.test(dat_groups$body_mass_g
)
```


### Question 14
Based on the results of the Bartlett tests, do you anticipate any issues with heterogeneity in any of the models? Make sure you justify your response with the results of your tests.

Since the p-value is 0.1741 it is pretty high and does not seem to be very significant. Because of this we will accept the null that that there is homogeneity. In previous questions (12 & 13), the p-value was significant so we would reject the null.  

