---
title: "Lab 7: Bootstrapping"
author: "Desiree Smith"
date: "10/27/2021"
output: html_document
---


```{r Walkthrough}
dat = matrix(1:49, nrow = 7, ncol = 7)
print(dat)

apply(dat, MARGIN = 1, FUN = min)

apply(dat, MARGIN = 1, FUN = max)

apply(dat, MARGIN = 2, FUN = mean)

require("here")
moths = read.csv(here("data", "moths.csv"))
head(moths)

b = 10000

# numeric() creates an vector of length m with all values initialized to zero
result = numeric(b)
head(result)

for(i in 1:b)
{
  result[i] = mean(sample(moths$anst, replace=TRUE))
}

mean(result) #2.47

quantile(result,c(0.025,0.975))

install.packages("boot")
require(boot)
boot(data, statistic, R)

boot_mean = function(x, i)
{
  return(mean(x[i], na.rm = TRUE))
}

myboot = 
  boot(
    data = moths$anst,
    statistic = boot_mean,
    R = 10000)
print(myboot)

str(myboot)

#Calculating mean and sd
mean(moths$anst)
myboot$t0
mean(myboot$t) - myboot$t0
sd(myboot$t)

#CI
quantile(
  myboot$t,
  c(0.025, 0.975))
moth_dat = moths[,-1] # how to remove a column
head(moth_dat)

```


```{r Walkthrough}
n = nrow(moth_dat) #number of rows or sample observations


m = 100 #number of bootstrap iterations
moth_result = matrix(
  nrow = m,
  ncol = n)
```


```{r Walkthrough}
n = nrow(moth_dat) #number of rows or sample observations

m = 100 #number of bootstrap iterations

moth_result = matrix(
  nrow = m,
  ncol = n)
```

Running the bootstrap simulation
```{r Running the bootstrap simulation}
n = nrow(moth_dat) #number of rows or sample observations

m = 100 #number of bootstrap iterations

moth_result = matrix(
  nrow = m,
  ncol = n)



# The outer loop: runs once for each bootstrap iteration.  index variable is i
for(i in 1:m)
{
  # The inner loop: simulates increasing sampling intensity
  # Sampling intensity ranges from 1 site to the complete count of sites (24)
  # index variable is j
  for(j in 1:n)
  {
    # sample the input data row indices, with replacement
    rows_j = sample(n, size = j, replace=TRUE)

    # Creates a new data matrix from the resampled rows.
    t1 = moth_dat[rows_j, ]

    # Calculates the column sums of the new data matrix.
    t2 = apply(t1, 2, sum)

    # Counts the number of columns in which any moths were observed
    moth_result[i, j] = sum(t2 > 0)
  }
}

```

First Draft
```{r First Draft}
rarefaction_sampler = function(input_dat, n_iterations)
{
  n = nrow(moth_dat) #number of rows or sample observations
  m = 100 #number of bootstrap iterations

  moth_result = matrix(
    nrow = m,
    ncol = n)

  # The outer loop: runs once for each bootstrap iteration.  index variable is i
  for(i in 1:m)
  {
    # The inner loop: simulates increasing sampling intensity
    # Sampling intensity ranges from 1 site to the complete count of sites (24)
    # index variable is j
    for(j in 1:n)
    {

      # sample the input data row indices, with replacement
      rows_j = sample(n, size = j, replace=TRUE)

      # Creates a new data matrix
      t1 = moth_dat[rows_j, ]

      # Calculates the column sums
      t2 = apply(t1, 2, sum)

      # Counts the number of columns in which any moths were observed
      moth_result[i, j] = sum(t2 > 0)
    }
  }

  return(moth_result)
}

rarefact = rarefaction_sampler(moth_dat, 100)

```
Second Draft
```{r Second Draft}
rarefaction_sampler = function(input_dat, n_iterations)
{
  n_input_rows = nrow(input_dat)

  results_out = matrix(
    nrow = n_iterations,
    ncol = n_input_rows)

  # The outer loop: runs once for each bootstrap iteration.  index variable is i
  for(i in 1:n_iterations)
  {
    # The inner loop: simulates increasing sampling intensity
    # Sampling intensity ranges from 1 site to the complete count of
    # sites in the input data (n)
    # index variable is j
    for(j in 1:n)
    {
      # sample the input data row indices, with replacement
      rows_j = sample(n, size = j, replace=TRUE)

      # Creates a new data matrix
      t1 = input_dat[rows_j, ]

      # Calculates the column sums
      t2 = apply(t1, 2, sum)

      # Counts the number of columns in which any moths were observed
      results_out[i, j] = sum(t2 > 0)
    }
  }
  return(results_out)
}

rarefact = rarefaction_sampler(moth_dat, 100)
head(rarefact)

```
Check in a fresh environment
```{r Check in a fresh environment}
 #This clears the current R session's environment
rm(list = ls())

# Re-read my data:
moths = read.csv(here("data", "moths.csv"))
moth_dat = moths[,-1]

rarefaction_sampler = function(input_dat, n_iterations,n_input_rows)
{
  n_input_rows = nrow(input_dat)

  results_out = matrix(
    nrow = n_iterations,
    ncol = n_input_rows)

  # The outer loop: runs once for each bootstrap iteration.  index variable is i
  for(i in 1:n_iterations)
  {
    # The inner loop: simulates increasing sampling intensity
    # Sampling intensity ranges from 1 site to the complete count of
    # sites in the input data (n)
    for(j in 1:n_input_rows)
    {
      # sample the input data row indices, with replacement
      rows_j = sample(n_input_rows, size = j, replace=TRUE)

      # Creates a new data matrix
      t1 = input_dat[rows_j, ]

      # Calculates the column sums
      t2 = apply(t1, 2, sum)

      # Counts the number of columns in which any moths were observed
      results_out[i, j] = sum(t2 > 0)
    }
  }
  return(results_out)
}

rarefact = rarefaction_sampler(moth_dat, 100)


head(rarefact)
```
Building the rarefaction curve
```{r Building the rarefaction curve}
moths = read.csv(here("data", "moths.csv"))
rarefact = rarefaction_sampler(moths[,-1], 10000)
rarefact
```
Plotting 
```{r Plotting}
rare_mean = apply(rarefact, 2, mean)
rare_quant = apply(rarefact, 2, quantile, probs=c(0.025, 0.975))
rare = t(rbind(rare_mean, rare_quant))

matplot(
  rare,
  type='l',
  xlab='Number of sampling plots',
  ylab='Species richness',
  main='Rarefaction Curve')

legend(
  'bottomright',
  legend=c('mean','2.5%','97.5%'),
  lty=c(1,2,3),col=c(1,2,3), inset=c(.1,.1))
```
__Question 1__
What is the sample size, n? 

```{r Q1}
require(palmerpenguins)
head(penguins)

 penguins$species == "Gentoo"


gen = subset(penguins,species == "Gentoo" )

sum.gen = sum(is.na(gen$bill_length_mm))

sum.gen

#The sample size is 124 but there is a NA so it will be 123. 

```
 __Question 2__
 What is the sample standard deviation?
 
```{r Q2}
sd(gen$bill_length_mm,na.rm = TRUE)

```
 
__Question 3__
What are the critical t-values?
```{r Q3}
crit = qt(0.95, 123,lower.tail = TRUE)#1.657

qt(0.95, 123,lower.tail = FALSE)

#The critical t- values are +-1.657

```

__Question 4__
What is the sample standard error?

```{r Q4}
rm(list = ls())

sse_mean = function(x)
  {na.x = is.na(x)
  x2 = x[!na.x]
  sd.x2 = sd(x2)
  n = length(x2)
  sse = sd.x2/(sqrt(n)) 
  return(sse)
}

gen = subset(penguins,species == "Gentoo" )
 sse = sse_mean(gen$bill_length_mm)

 
 sse


```

__Question 5__
Construct the CI and show the R code you used for the calculation.

```{r Q5}
(crit * sse)
#0.461

#The parametric CI is 0.278 +- 0.461

```

__Question 6 - 8__


```{r Q6 - Q8 {.tabset. tabset-pills}


require(boot)
boot(data, statistic, R)

boot_mean = function(x, i)
{
  return(mean(x[i], na.rm = TRUE))
}

mypenboot = 
  boot(
    data = gen$bill_length_mm,
    statistic = boot_mean,
    R = 10000)
print(mypenboot)

str(mypenboot)

#Calculating mean and sd
mean(gen$bill_length_mm, na.rm = TRUE)# the mean is 47.40

mypenboot$t0

mean(mypenboot$t) - mypenboot$t0

sd(mypenboot$t)

#CI
quantile(
  mypenboot$t,
  c(0.025, 0.975))


```

