---
title: "Lab3:DataExploration_D.Smith"
subtitle: "Desiree Smith
author: "Worked with Alex Fink"
date: "9/22/2021"
output: html_document
---

```{r}
install.packages("psych")
require(psych)
pairs.panels(iris)


```
```{r}
require(here)
dat_bird = read.csv(here("data", "bird.sta.csv"))
head(dat_bird)
dat_habitat = data.frame(read.csv(here("data", "hab.sta.csv")))
dat_habitat
dat_all = merge(dat_bird,data_habitat)
dat_all

plot(ba.tot ~ elev, data = dat_all)
```

```{r}
sample(dat_all$CEWA, 100)

my_vec = rep(1:3, 5)
my_vec == 3
my_vec > 1

cewa = dat_all$CEWA 
cewa

cewa_present_absent = (cewa >= 1)
cewa_present_absent

plot(x = dat_all$elev, y = cewa_present_absent)

# Function to calculate the logistic parameter a given the slope and midpoint
get_logistic_param_a = function(slope, midpoint)
{
  b = slope / 4
  return (-midpoint * (slope / 4))
}

# Function to calculate the logistic parameter b given the slope
get_logistic_param_b = function(slope)
{
  return (slope / 4)
}


# Calculate the value of the logistic function at x, given the parameters a and b.
logistic = function(x, a, b)
{
  val = exp(a + b * x)
  return(val / (1 + val))
}

# Calculate the value of the logistic function at x, given a slope and midpoint.
logistic_midpoint_slope = function(x, midpoint, slope)
{
  b = get_logistic_param_b(slope)
  a = get_logistic_param_a(slope, midpoint)
  return(logistic(x, a, b))
}

#Positive slope
{
plot(x = dat_all$elev, y = cewa_present_absent)
curve(logistic_midpoint_slope(x, midpoint = 370, slope = 0.05), add = TRUE)
}

#negative slope
plot(x = dat_all$elev, y = cewa_present_absent)
curve(logistic_midpoint_slope(x, midpoint = 400, slope = -0.1), add = TRUE)

#Shallower negative slope
plot(x = dat_all$elev, y = cewa_present_absent)
curve(logistic_midpoint_slope(x, midpoint = 400, slope = -0.05), add = TRUE)




```
 
__Question 1__
The basal area is square meters per hectares.

__Question 2__
```{r}
require(psych)
pairs.panels(dat_all[ , c("elev", "ba.tot", "slope", "aspect")])

```

__Question 3__

```{r}
WAVI = dat_all$WAVI
WAVI

wavi_present_absent = (WAVI >= 1)
wavi_present_absent

# Function to calculate the logistic parameter a given the slope and midpoint
get_logistic_param_a = function(slope, midpoint)
{
  b = slope / 4
  return (-midpoint * (slope / 4))
}

# Function to calculate the logistic parameter b given the slope
get_logistic_param_b = function(slope)
{
  return (slope / 4)
}


# Calculate the value of the logistic function at x, given the parameters a and b.
logistic = function(x, a, b)
{
  val = exp(a + b * x)
  return(val / (1 + val))
}

# Calculate the value of the logistic function at x, given a slope and midpoint.
logistic_midpoint_slope = function(x, midpoint, slope)
{
  b = get_logistic_param_b(slope)
  a = get_logistic_param_a(slope, midpoint)
  return(logistic(x, a, b))
}

plot(x = dat_all$ba.tot, y = wavi_present_absent, col = "orange")
curve(logistic_midpoint_slope(x, midpoint = 100, slope = -0.6), add = TRUE)

```
__Question 4__
The birds seem to prefer areas with low tree cover. They were found in area with lower tree densities. There are also a couple birds that were found in an area with high tree cover but the majority of them that are present are in lower tree densities. 

__Question 5__
```{r}
SWTH = dat_all$SWTH
SWTH

swth_present_absent = (SWTH >= 1)
swth_present_absent

# Function to calculate the logistic parameter a given the slope and midpoint
get_logistic_param_a = function(slope, midpoint)
{
  b = slope / 4
  return (-midpoint * (slope / 4))
}

# Function to calculate the logistic parameter b given the slope
get_logistic_param_b = function(slope)
{
  return (slope / 4)
}


# Calculate the value of the logistic function at x, given the parameters a and b.
logistic = function(x, a, b)
{
  val = exp(a + b * x)
  return(val / (1 + val))
}

# Calculate the value of the logistic function at x, given a slope and midpoint.
logistic_midpoint_slope = function(x, midpoint, slope)
{
  b = get_logistic_param_b(slope)
  a = get_logistic_param_a(slope, midpoint)
  return(logistic(x, a, b))
}

plot(x = dat_all$ba.tot, y = swth_present_absent, col = "blue")
curve(logistic_midpoint_slope(x, midpoint = 100, slope = -0.7), add = TRUE)
```

__Question 6__
The birds seem to prefer areas with low tree cover. On the plot there is a lot of birds present in areas with low tree cover on the left top hand corner of the graph. 

__Question 7 and Question 8__

```{r}
total_GRJA = dat_all$GRJA
sum(total_GRJA)

#The total number of Gray Jays that were observed were 181
```

__Question 9 and Question 10__

```{r}
my_vec = dat_all$GRJA
my_vec >= 1
sum_vec = as.numeric(my_vec)
sum(my_vec)

my_vec


grja_present_absent = (my_vec >= 1)
grja_present_absent
sum(grja_present_absent)

#The total number of sample sites that the Gray Jays were observed were 110
```

```{r}
install.packages("yaml")
require("yaml")
```

