---
title: "Lab4"
subtitle: Desiree Smith
author: Alex Fink, JT and Ethan
date: "9/29/2021"
output:
  html_document: default
  pdf_document: default
---


```{r}
x = seq(-3, 3, length.out = 1000)
y = dnorm(x)

plot(x, y, main = "Normal PDF", type = "l")
abline(h = 0)

```
```{r}
require(palmerpenguins)
hist(
  penguins$body_mass_g,
  main = "Histogram of Penguin Body Mass",
  xlab = "Body Mass (g)"
)

mean(penguins$body_mass_g, na.rm = TRUE) #4201.754
sd(penguins$body_mass_g, na.rm = TRUE) #801.95
nrow(penguins) #344
```

```{r}
dat_1 = rnorm(n = 344, mean = 202, sd = 802)
dat_2 = rnorm(n = 344, mean = 4202, sd = 802)
dat_3 = rnorm(n = 344, mean = 4202, sd = 802)
dat_4 = rnorm(n = 344, mean = 4202, sd = 802)

par(mfrow = c(2,2))

hist(dat_1)
hist(dat_2)
hist(dat_3)
hist(dat_4)
```

```{r}
set.seed(12)
dat_unif = runif(n =27, min = 0, max = 4)
hist(dat_unif)

set.seed(1)
dat_unif_1 = runif(n = 270, min = 0, max = 4)
set.seed(1)
dat_unif_2 = runif(n = 270, min = 0, max = 4)

par(mfrow = c(1, 2))
hist(dat_unif_1)
hist(dat_unif_2)

```

```{r}
set.seed(123)
n = 17
slope = 0.7
intcp = 0.2

guess_x = 6
guess_y = 4
guess_slope = 0.72

x = runif(n = n, min = 1, max = 10)
y = rnorm(n = n, mean = slope * x + intcp)

line_point_slope = function(x, x1, y1, slope)
{
  get_y_intercept = 
    function(x1, y1, slope) 
      return(-(x1 * slope) + y1)
  
  linear = 
    function(x, yint, slope) 
      return(yint + x * slope)
  
  return(linear(x, get_y_intercept(x1, y1, slope), slope))
}

plot(x, y, pch = 16)
curve(line_point_slope(x, guess_x, guess_y, guess_slope), add = T)

n_pts = 10
x_min = 1
x_max = 10
x = runif(n = n_pts, min = x_min, max = x_max)

dat = data.frame(x = x, y_observed = rnorm(n_pts))

plot(y_observed ~ x, data = dat, pch = 8)

line_point_slope = function(x, x1, y1, slope)
{
  get_y_intercept = 
    function(x1, y1, slope) 
      return(-(x1 * slope) + y1)
  
  linear = 
    function(x, yint, slope) 
      return(yint + x * slope)
  
  return(linear(x, get_y_intercept(x1, y1, slope), slope))
  
}

n_pts = 10
x_min = 1
x_max = 10
x = runif(n = n_pts, min = x_min, max = x_max)

dat = data.frame(x = x, y_observed = rnorm(n_pts))

plot(y_observed ~ x, data = dat, pch = 8)

guess_x = 6
guess_y = 0
guess_slope = 0.1

plot(y_observed ~ x, data = dat, pch = 8)
curve(line_point_slope(x, guess_x, guess_y, guess_slope), add = T)

dat$y_predicted = line_point_slope(dat$x, guess_x, guess_y, guess_slope)


dat$resids = (dat$y_observed - dat$y_predicted)

sum(dat$resids)


```


__Question 1__
```{r}
norm_17 = rnorm(n = 17, mean = 10.4, sd = 2.4)
norm_30 = rnorm(n = 30, mean = 10.4, sd = 2.4)
norm_300 = rnorm(n = 300, mean = 10.4, sd = 2.4)
norm_3000 = rnorm(n = 3000, mean = 10.4, sd = 2.4)
```

__Question 2__
```{r}
par(mfrow = c(2,2))

hist(norm_17, main = "17 Random Data Points", xlab = "Data Points", col = "rosybrown")
hist(norm_30,  main = "30 Random Data Points", xlab = "Data Points", col = "seashell")
hist(norm_300,  main = "300 Random Data Points", xlab = "Data Points", col ="lightcyan")
hist(norm_3000,  main = "3000 Random Data Points", xlab = "Data Points", col = "palegreen")


```
__Question 3__
```{r}
image_file = "lab_04_hist_01.0.png"

require(here)

{
png(
  here("Images", "lab_04_hist_01.png"),
  width = 1500, height = 1600)
par(mfrow = c(2,2))

hist(norm_17, main = "17 Random Data Points", xlab = "Data Points", col = "rosybrown")
hist(norm_30,  main = "30 Random Data Points", xlab = "Data Points", col = "seashell")
hist(norm_300,  main = "300 Random Data Points", xlab = "Data Points", col ="lightcyan")
hist(norm_3000,  main = "3000 Random Data Points", xlab = "Data Points", col = "palegreen")

dev.off()
}


```


__Question 4__
Two of the histogram have a normal distribution. They have a bell shaped curve.Two seem to be more uniform. There is not a distinct curve. 

__Question 5__
The shapes of the histograms are different because I am using rnorm.This will give me random numbers each time. It seems that as the data points increases the shape started to become more like a bell curve. 

__Question 6__
The parameters that I used for the Normal distribution are mean and standard deviation.

__Question 7__
```{r}

require(here)

pdf(file = here("Images", "norm_1.pdf"))

x = seq(3, 19, length.out = 1000)
y = dnorm(x, mean = 10.4, sd = 2.4)

plot(x, y, main = "Standard Normal PDF: mean = 10.4 and sd = 2.4", type = "l", xlim = c(3, 19))
abline(h = 0)






```

__Question 8__
Upload File

__Question 9__
```{r}
n_pts = 56
x_min = 1
x_max = 80
x = runif(n = n_pts, min = x_min, max = x_max)
dat_4 = data.frame(x = x, y_observed = rnorm(n_pts))
dat_4

rnorm(56)
set.seed(123)

par(mfrow = c(2,2))

image_file = "Random_Plots.png"

require(here)

{
png(
  here("Images", "Random_Plots.png"),
  width = 1500, height = 1600)
par(mfrow = c(2,2))

plot(dat_4, main = "Random Numbers", xlab = "data points", ylab = "y observed", col = "palegreen4")

boxplot(dat_4, main = "Random Number 2.0", xlab = "data points", ylab = "y observed", col = "red4")

hist(dat_4$x, main = "Random Number 3.0", xlab = "data points", ylab = "y observed", col = "lavenderblush",
border = "plum4")

barplot(dat_4$y_observed, main = "Random Number 4.0", xlab = "data points", ylab = "y observed", col = "darkblue")

}


```

__Question 10__
Upload image 

__Question 11__
```{r}

png(
  here("Images", "Line_Slope.png"),
  width = 1500, height = 1600)

line_point_slope = function(x, x1, y1, slope)
  
{
  get_y_intercept = 
    function(x1, y1, slope) 
      return(-(x1 * slope) + y1)
  
  linear = 
    function(x, yint, slope) 
      return(yint + x * slope)
  
  return(linear(x, get_y_intercept(x1, y1, slope), slope))
  
}

plot(dat_4, main = "Random Numbers", xlab = "data points", ylab = "y observed", col = "palegreen4")

logistic_midpoint_slope = function(x, midpoint, slope)
{
  b = get_logistic_param_b(slope)
  a = get_logistic_param_a(slope, midpoint)
  return(logistic(x, a, b))
}

guess_x = 5
guess_y = -3
guess_slope = 0.1

plot(y_observed ~ x, data = dat_4, pch = 8, col = "darkmagenta")
curve(line_point_slope(x, guess_x, guess_y, guess_slope), add = T)


```
__Question 12__
Upload image

__Question 13 and Question 14__
```{r}

dat_4$y_predicted = line_point_slope(dat_4$x, guess_x, guess_y, guess_slope)
dat_4


dat_4$resids = (dat_4$y_observed - dat_4$y_predicted)

dat_4

#Question 14

hist(dat_4$resids, main = "Model Residuals", xlab = "data points", ylab = "y observed", col = "darkorange3",
border = "gold2")

plot(dat_4$y_predicted, dat_4$resids, main = "Model's Predicted Values", xlab = "Predicted Values", ylab = "Residuals", col = "indianred") 

 

```

